\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage{enumitem}
\title{User Stories}
\date{}

\begin{document}

\maketitle

\section{Authentication}
\begin{itemize}[leftmargin=*, label=--]
    \item As a user, I want to log in securely, so that I can access my personalized dashboard and data.
    \item As an admin, I want to reset the user database, so that I can ensure only authorized users have access.
\end{itemize}

\section{Project Management}
\begin{itemize}[leftmargin=*, label=--]
    \item As a user, I want to create new projects, so that I can organize my work and tasks efficiently.
    \item As a user, I want to view a list of all my projects, so that I can quickly access and manage them.
    \item As a user, I want to edit project details such as title, description, and deadline, so that project information stays up to date.
    \item As a user, I want to delete projects, so that I can remove obsolete or completed projects from my workspace.
\end{itemize}

\section{Team and Role Management}
\begin{itemize}[leftmargin=*, label=--]
    \item As a project owner, I want to add team members to a project and assign roles, so that I can collaborate with others.
    \item As a project owner, I want to remove team members, so that I can manage project access and responsibilities.
    \item As a project owner, I want to assign or change the project leader, so that leadership can be transferred as needed.
    \item As a team member, I want to view the list of project members and their roles, so that I know who is involved.
\end{itemize}

\section{Task and Subtask Management}
\begin{itemize}[leftmargin=*, label=--]
    \item As a user, I want to add tasks and subtasks to projects, so that I can break down work into manageable pieces.
    \item As a user, I want to assign tasks and subtasks to team members, so that responsibilities are clear.
    \item As a user, I want to edit and delete tasks and subtasks, so that I can keep project plans accurate.
    \item As a user, I want to set deadlines, durations, and dependencies for tasks and subtasks, so that project timelines are well-defined.
    \item As a user, I want to categorize tasks and subtasks using the Eisenhower matrix, so that I can prioritize my work effectively.
    \item As a user, I want to drag and drop tasks/subtasks between categories, so that I can easily update priorities.
\end{itemize}

\section{Visualization and Planning}
\begin{itemize}[leftmargin=*, label=--]
    \item As a user, I want to view my tasks and subtasks in a dashboard, so that I can see what needs to be done today.
    \item As a user, I want to see a Gantt chart of project tasks and subtasks, so that I can visualize timelines and dependencies.
    \item As a user, I want to view a calendar with deadlines, public holidays, and personal time off, so that I can plan my work schedule.
\end{itemize}

\section{Event Logging}
\begin{itemize}[leftmargin=*, label=--]
    \item As a user, I want to see an event log of actions taken in the system, so that I can track project history and changes.
    \item As a user, I want to log significant actions (e.g., login, project changes, team updates), so that there is an audit trail.
\end{itemize}

\section{AI/LLM Integration}
\begin{itemize}[leftmargin=*, label=--]
    \item As a user, I want to receive AI-powered suggestions for task categorization, so that I can prioritize tasks more effectively.
    \item As a user, I want the system to generate commit summaries for project file changes using an LLM, so that version control messages are meaningful.
    \item As a user, I want to interact with local LLMs (TinyLlama, DeepSeek) for planning and suggestions, so that I can leverage AI capabilities offline.
\end{itemize}

\section{Miscellaneous}
\begin{itemize}[leftmargin=*, label=--]
    \item As a user, I want to manage my personal settings, so that I can customize my experience.
    \item As a user, I want to log out securely, so that my account remains protected.
\end{itemize}

\end{document}